---
title: "Forecasting Competition -- Getting Started"
author: "Lee Spitzley"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)

```

# Overview
This document contains some basic intructions to get you started with the 
forecasting competition.

You will want the `GSODR`, `ggplot2`, `lubridate` libraries to run this lab.

```{r packages, message=FALSE}
library(GSODR)
library(ggplot2)
library(lubridate)
```
Also, include the helpful forecast functions I made that will help you format your forecasts.
```{r}
source('forecast_functions.R')
```
# Data
You can download the data from GSOD using the format below.
```{r download, eval=FALSE}
last_ten_years <-  get_GSOD(years = 2009:2019, station = '725180-14735')
saveRDS(object = last_ten_years, file = 'last_ten_alb.rds')
```

It is often faster to just download all of the data once,
and then append the newest data to the end of the dataframe.

```{r append, eval=FALSE}
last_ten_years <- readRDS('last_ten_alb.rds')
current_year <- get_GSOD(years = 2019, station = '725180-14735')
new_rows <- subset(current_year, !(YEARMODA %in% last_ten_years$YEARMODA))
last_ten_years <- rbind(last_ten_years, new_rows)
last_ten_years$I_PRCP <- ifelse(last_ten_years$PRCP > 0, 1, 0)
saveRDS(object = last_ten_years, file = 'last_ten_alb.rds')
```

Storing data in the `rds` format preserves column data types and other metadata.

```{r read_cached, cache=TRUE, echo=FALSE}
last_ten_years = readRDS('last_ten_alb.rds')
```

Show the data for the last ten years.
```{r visual, cache=FALSE, warning=FALSE, message=FALSE}
qplot(x = last_ten_years$YEARMODA, y = last_ten_years$MAX)

yoy <- ggplot(last_ten_years, aes(x=YDAY, y=MAX, color=YEAR))
yoy + geom_point() + geom_smooth()
```

# Starting Your Forecast
Get the dates that you want to predict `eval.dates` and the previous 
week `start_prev_seven`, which is one way to create your predictions.
```{r get_dates, echo=1:4}
prediction_start <- Sys.Date() + days(1)
start_prev_seven <- prediction_start - days(7)
eval.dates <- seq(prediction_start, prediction_start + 6, by='days')
eval_df_base <- as.data.frame(eval.dates)
prediction_start <- date('2019-02-03') + days(1)
start_prev_seven <- prediction_start - days(7)
eval.dates <- seq(prediction_start, prediction_start + 6, by='days')
eval_df_base <- as.data.frame(eval.dates)
kable(eval_df_base)
```

# Example Forecast
Predict the future temperature as the average of the last seven days.

```{r prev_seven_mean, echo=1:11}
forecast_name <- 'mean_prev_seven'
prev_seven <- subset(last_ten_years, YEARMODA >= start_prev_seven & 
                       YEARMODA < prediction_start)
mean.max <- mean(prev_seven$MAX) 
mean.min <- mean(prev_seven$MIN)
mean.prcp <- mean(prev_seven$I_PRCP)

pred.df.list <- list()
pred.df.list[['max']] <- forecast_df_builder(eval.dates, 'mean_prev_seven', 'MAX', 
                                             rep(mean.max, 7))
pred.df.list[['min']] <- forecast_df_builder(eval.dates, 'mean_prev_seven', 
                                             'MIN', rep(mean.min, 7))
pred.df.list[['i_prcp']] <- forecast_df_builder(eval.dates, 'mean_prev_seven', 
                                             'I_PRCP', rep(mean.prcp, 7))
pred.df <- do.call(rbind, pred.df.list)
kable(pred.df, digits = 4, caption = "Predicted max, min, and precipitation")
```

